---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
```

```{r Loading and preprocessing the data}
# Set dir
setwd("/home/carlos/coursera/R/Reproducible_Research/week02/")

# Download file and unzip
#temp <- tempfile()
#download.file("https://github.com/ceggithub/RepData_PeerAssessment1/raw/master/activity.zip", temp)
#unzip(temp)
#unlink(temp)

# Load data
activity <- read.csv2(file = "activity.csv", header = TRUE, sep = ',', na.strings = "NA" )
activity$date <- strptime(activity$date, format = "%Y-%m-%d")

# Sum, Mean and Median steps by day
byday <- ddply(activity,~date,summarise,sum=sum(steps), mean=mean(steps, na.rm = TRUE), median=median(steps, na.rm = TRUE))

# Mean and Median steps by time series
byfiveminutes <- ddply(activity, ~interval, summarise, mean=mean(steps, na.rm = TRUE), median=median(steps, na.rm = TRUE))

```

## What is mean total number of steps taken per day?
```{r, echo=FALSE}
# Steps frequency by 5-minutes
hist(byday$sum, xlab = "Total Steps by day", main = "Frequency Steps", col = "red")
```

### Results by day
```{r, echo=FALSE}
byday
```


## What is the average daily activity pattern?
```{r, include=FALSE}
# Maximum average by 5-minutes
maximo <- byfiveminutes[byfiveminutes$mean==max(byfiveminutes$mean),c('interval')]
```
Average maximum occurs within: `r maximo`

```{r, echo=FALSE}
plot(byfiveminutes$interval, byfiveminutes$mean, col="red", type = "b", main = "Mean steps by 5-minute", ylab = "Steps", xlab = "Minute")

```



## Imputing missing values

Missing are `r sum(is.na(activity$steps))` rows

This missing data will are filled by median of the 5-minute period.


### The new data look

```{r, echo=FALSE}
# Create new data frame and fill missing steps with median
activityfilled <- activity
activityfilled <- within(activity, steps <- ifelse(is.na(steps), byfiveminutes$median, steps))

# Re-make by day
bydayfilled <- ddply(activityfilled,~date,summarise,sum=sum(steps),mean=mean(steps, na.rm = TRUE), median=median(steps, na.rm = TRUE))

# Re-make by 5-minutes
byfiveminutesfilled <- ddply(activityfilled, ~interval, summarise, mean=mean(steps, na.rm = TRUE), median=median(steps, na.rm = TRUE))

# Plot results
hist(bydayfilled$sum, xlab = "Total Steps by day", main = "Frequency Steps filled", col = "red")
```

Results by day with missed filled

How to fill the average of the values it was high

```{r, echo=FALSE}
bydayfilled
```


## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=FALSE}
# Create new var: dayweek
activity <- within(activity, dayweek <- as.factor(ifelse((weekdays(date) == "Saturday" | weekdays(date) == "Sunday"), "Weekend", "Weekday")))

# Re-make by dayweek
byfiveminutesweekends <- ddply(activity[activity$dayweek=="Weekend", ],~interval,summarise,mean=mean(steps, na.rm = TRUE))
byfiveminutesweekdays <- ddply(activity[activity$dayweek=="Weekday", ],~interval,summarise,mean=mean(steps, na.rm = TRUE))

# Plot graphs
par(mfrow=c(2,1))
plot(byfiveminutesweekdays$interval, byfiveminutesweekdays$mean, col="red", type = "l", main = "Weekday Steps by 5-minute", ylab = "Steps", xlab = "", ylim = range(byfiveminutesweekdays$mean))
plot(byfiveminutesweekends$interval, byfiveminutesweekends$mean, col="blue", type = "l", main = "Weekend Steps by 5-minute", ylab = "Steps", xlab = "", ylim = range(byfiveminutesweekdays$mean))

```

